Practical Machine Learning Course Project 
========================================================

The goal of this project is to predict the manner in which they did the exercise. Due to the limit of the available time and new to the RStudio, I have decided to test RandomForest algorithm first, and will add and compare with other algorithms later. 

Load required libraries and set seed for reproducibility

```{r}
library(caret)
library(rpart)
library(randomForest)
set.seed(1234)
```

Read data files and make training and validation data sets.
(1) Assume the data files are in current working directory.
(2) 60% for training and 40% for validation - since the data is big enough.

```{r}
rm(list=ls())
fileTraining <- "C:/Users/TeddyKo/R_working/pml-training.csv"
trainingCSV = read.csv(fileTraining,na.strings=c("NA",""), header=TRUE)
inTrain <- createDataPartition(trainingCSV$classe, p=0.60,list=FALSE)
training <- trainingCSV[inTrain, ]
validation <- trainingCSV[-inTrain, ]
```

Create summary and view the top few records

```{r}
## summary(training)
## head(training)
```

Some of the columns have no data, it might not be productive to use them as-is. Therefore filtering out fields with a lot of (more than 60%) null values.

```{r}
goodVars <- c((colSums(is.na(training[,-160])) >= 0.4*nrow(training)),160)
training <- training[,goodVars]
dim(training)
validation<-validation[,goodVars]
dim(validation)
training<-training[complete.cases(training),]
dim(training)
```

Train the model using RandomForest on the training data set.

```{r}
model <- randomForest(classe ~ ., data=training)
print(model)
```

Evaluate the model on the evaluation dataset.

```{r}
confusionMatrix(predict(model,newdata=validation[,-ncol(validation)]), validation$classe)
accurate <- c(as.numeric(predict(model,newdata=validation[,-ncol(validation)])==validation$classe))
accuracy <- sum(accurate)*100/nrow(validation)
message("Model Accuracy using Validation set = " , format(round(accuracy, 2), nsmall = 2), "%")
```

Predict the class values in the testing set with 20 observations

```{r}
fileTesting <- "C:/Users/TeddyKo/R_working/pml-testing.csv"
testing = read.csv(fileTesting, na.strings=c("NA",""), header=TRUE)
dim(testing)
testing <- testing[,goodVars]
dim(testing)
predictions <- predict(model,newdata=testing)
predictions
```

